---
title: "caphistory2"
author: "Bethany Liberto"
date: "June 16, 2019"
output:
  word_document: default
editor_options: 
  chunk_output_type: console
---

Using CJS and JS. JS -> for population estimates. CJS -> for time-dependent survival and recapture rates of a given species 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      results = "hide") ##This just hides outputs when knitting because it was like 60 pages
options(warn=-1) ##This hides warnings, again, just really for knitting. Can be turned back on with warn=0
```
  
  
```{r}
library(marked)
library(RMark)
library(ggplot2)
```


abundance/effort AIC vvvv

```{r}

data <- read.csv("Dragonfly1718Abundance.csv", stringsAsFactors = F)

### Each possible combination of variables needs to be analyzed.  The lowest score (not counting the full model?) wins. #####
mod1<-glm(data$adjusted.abundance~data$pH)
mod2<-glm(data$adjusted.abundance~data$elevation)
mod3<-glm(data$adjusted.abundance~data$fishpresence)
mod4<-glm(data$adjusted.abundance~data$lenticlotic)  
mod5<-glm(data$adjusted.abundance~data$openwater)
mod6<-glm(data$adjusted.abundance~data$pH*data$elevation)
mod7<-glm(data$adjusted.abundance~data$pH*data$elevation*data$fishpresence)
mod8<-glm(data$adjusted.abundance~data$pH*data$elevation*data$fishpresence*data$lenticlotic)
mod9<-glm(data$adjusted.abundance~data$pH*data$elevation*data$fishpresence*data$lenticlotic*data$openwater) 
mod10<-glm(data$adjusted.abundance~data$elevation*data$fishpresence)
mod11<-glm(data$adjusted.abundance~data$elevation*data$fishpresence*data$lenticlotic)
mod12<-glm(data$adjusted.abundance~data$elevation*data$fishpresence*data$lenticlotic*data$openwater)
mod13<-glm(data$adjusted.abundance~data$fishpresence*data$lenticlotic)
mod14<-glm(data$adjusted.abundance~data$fishpresence*data$openwater) #### Chicken dinner  #####
mod15<-glm(data$adjusted.abundance~data$fishpresence*data$lenticlotic*data$openwater) #### Almost the same ####
mod16<-glm(data$adjusted.abundance~data$lenticlotic*data$openwater)



AIC(mod1, k=2)  ### k = 2 is the "penalty" for adding more factors #####
AIC(mod2, k=2)
AIC(mod3, k=2)
AIC(mod4, k=2) 
AIC(mod5, k=2)
AIC(mod6, k=2)
AIC(mod7, k=2)
AIC(mod8, k=2)
AIC(mod9, k=2) 
AIC(mod10, k=2)
AIC(mod11, k=2)
AIC(mod12, k=2)
AIC(mod13, k=2)
AIC(mod14, k=2) #### Chicken dinner  #####
AIC(mod15, k=2) #### Almost the same, 0.1258 difference ####
AIC(mod16, k=2)


```




richness (just to see) vvvvv

```{r}
### Each possible combination of variables needs to be analyzed.  The lowest score (not counting the full model?) wins. #####
mod1<-glm(data$Richness~data$pH)
mod2<-glm(data$Richness~data$elevation)
mod3<-glm(data$Richness~data$fishpresence)
mod4<-glm(data$Richness~data$lenticlotic)  #### Chicken dinner  #####
mod5<-glm(data$Richness~data$openwater)
mod6<-glm(data$Richness~data$pH*data$elevation)
mod7<-glm(data$Richness~data$pH*data$elevation*data$fishpresence)
mod8<-glm(data$Richness~data$pH*data$elevation*data$fishpresence*data$lenticlotic)
mod9<-glm(data$Richness~data$pH*data$elevation*data$fishpresence*data$lenticlotic*data$openwater)
mod10<-glm(data$Richness~data$elevation*data$fishpresence)
mod11<-glm(data$Richness~data$elevation*data$fishpresence*data$lenticlotic)
mod12<-glm(data$Richness~data$elevation*data$fishpresence*data$lenticlotic*data$openwater)
mod13<-glm(data$Richness~data$fishpresence*data$lenticlotic)
mod14<-glm(data$Richness~data$fishpresence*data$openwater)
mod15<-glm(data$Richness~data$fishpresence*data$lenticlotic*data$openwater)
mod16<-glm(data$Richness~data$lenticlotic*data$openwater)



AIC(mod1, k=2)  ### k = 2 is the "penalty" for adding more factors #####
AIC(mod2, k=2)
AIC(mod3, k=2)
AIC(mod4, k=2) #### Chicken dinner  #####
AIC(mod5, k=2)
AIC(mod6, k=2)
AIC(mod7, k=2)
AIC(mod8, k=2)
AIC(mod9, k=2)
AIC(mod10, k=2)
AIC(mod11, k=2)
AIC(mod12, k=2)
AIC(mod13, k=2)
AIC(mod14, k=2)
AIC(mod15, k=2)
AIC(mod16, k=2)

```



these are all the JS and CJS for each location. Starting with NMBS
```{r}
capNMBS <- read.csv("capturehistoryNMBS.csv", stringsAsFactors = F)
str(capNMBS)
print(capNMBS)
summary(capNMBS)
capNMBS <- capNMBS[which(!is.na(capNMBS$c1)), ]
summary(capNMBS)

library(reshape)
## Format data for MARK

# get data formatted for MARK/RMark
capNMBS$ch <- paste(capNMBS$c1,
                capNMBS$c2,
                capNMBS$c3,
                capNMBS$c4,
                capNMBS$c5,
                capNMBS$c6,
                capNMBS$c7,
                capNMBS$c8,
                capNMBS$c9,
                capNMBS$c10,
                capNMBS$c11,
                capNMBS$c12,
                " ",
                ";",
                sep="")


data=data.frame(capNMBS)
data
transform=melt(capNMBS, id.vars="mark.number")
pivot=cast(transform, mark.number ~ value)
pivot[is.na(pivot)]=0

pivot[,2:ncol(pivot)][pivot[,2:ncol(pivot)] != 0] = 1
lh <- 11;


pivot$eh <- apply(pivot[2:lh],1,paste,collapse="") # concatenates encounter columns into eh
pivot[2:lh] <- NULL # drops individual encounter columns
# create commented tag
pivot$mark.number <- paste("/*", pivot$mark.number, "*/", sep=" ")
# sort by descending encounter histories
pivot <- pivot[order(data$ch,decreasing=TRUE),]
# tack on the frequency for the individual
pivot$end <- "1;";
# write out the input file
write.table(pivot,file="cjs-pivot.inp",sep=" ",quote=F,col.names=F,row.names=F);

pivot

```



```{r}
## Format data for marked package

marked_dfNMBS <- capNMBS

marked_dfNMBS$ch <- paste0(capNMBS$c1,
                capNMBS$c2,
                capNMBS$c3,
                capNMBS$c4,
                capNMBS$c5,
                capNMBS$c6,
                capNMBS$c7,
                capNMBS$c8,
                capNMBS$c9,
                capNMBS$c10,
                capNMBS$c11,
                capNMBS$c12)
                
model=crm(marked_dfNMBS, time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9))
model
model=cjs.hessian(model)
model
```

CJS: AIC  :  184.6449, Phi.(Intercept) -0.4564347, p.(Intercept)   -1.5498460

looking at sex and species. phi= apparent survival estimate. p=capture probability (percents) N=population estimate

```{r}
str(marked_dfNMBS)
marked_dfNMBS$sex.fac <- as.factor(marked_dfNMBS$sex)
marked_dfNMBS$species.fac <- as.factor(marked_dfNMBS$species)
modNMBS <- crm(marked_dfNMBS, 
        model="JS", 
        groups = c("sex.fac"), 
        model.parameters=list(Phi = list(formula = ~sex.fac), 
                              p = list(formula = ~1)))
modNMBS$results$beta #found online but not sure how to use it
modNMBS

cjs.hessian(modNMBS)

str(modNMBS)
# get values on real scale and not logit or log scales
modNMBS$results$reals


modNMBS <- crm(marked_dfNMBS, 
        model="JS", 
        groups = c("species.fac"), 
        model.parameters=list(Phi = list(formula = ~species.fac), 
                              p = list(formula = ~species.fac)))
modNMBS$results$beta 
modNMBS
modNMBS$results$reals
```


sex.fac : AIC  :  416.033, N estimate=  5.439211 species.fac: AIC  :  750.5, N estimate= 3.55226e-11

# Run with RMark (need to download Program MARK in order for this to work). This is looking at NMBS only

```{r}
MarkPath='C:/Program Files (x86)/MARK'
mark(marked_dfNMBS, model.parameters=list(p=list(formula=~1)),output=FALSE)$results$beta

mark(marked_dfNMBS, model = "Jolly", model.parameters=list(p=list(formula=~1)),output=FALSE)$results$beta
```

 AICc { Phi(~1)p(~1) } = 184.69827 , Phi:(Intercept) -0.4564347, p:(Intercept)   -1.5498461

####okay now for CT

```{r}
capCT <- read.csv("capturehistoryCT.csv", stringsAsFactors = F)
str(capCT)
print(capCT)
summary(capCT)
capCT <- capCT[which(!is.na(capCT$c1)), ]
summary(capCT)

library(reshape)
## Format data for MARK

# get data formatted for MARK/RMark
capCT$ch <- paste(capCT$c1,
                capCT$c2,
                capCT$c3,
                capCT$c4,
                capCT$c5,
                capCT$c6,
                capCT$c7,
                capCT$c8,
                capCT$c9,
                " ",
                ";",
                sep="")


data=data.frame(capCT)
data
transform=melt(capCT, id.vars="mark.number")
pivot=cast(transform, mark.number ~ value)
pivot[is.na(pivot)]=0

pivot[,2:ncol(pivot)][pivot[,2:ncol(pivot)] != 0] = 1
lh <- 11;


pivot$eh <- apply(pivot[2:lh],1,paste,collapse="") # concatenates encounter columns into eh
pivot[2:lh] <- NULL # drops individual encounter columns
# create commented tag
pivot$mark.number <- paste("/*", pivot$mark.number, "*/", sep=" ")
# sort by descending encounter histories
pivot <- pivot[order(data$ch,decreasing=TRUE),]
# tack on the frequency for the individual
pivot$end <- "1;";
# write out the input file
write.table(pivot,file="cjs-pivot.inp",sep=" ",quote=F,col.names=F,row.names=F);

pivot

```



```{r}
## Format data for marked package

marked_dfCT <- capCT

marked_dfCT$ch <- paste0(capCT$c1,
                capCT$c2,
                capCT$c3,
                capCT$c4,
                capCT$c5,
                capCT$c6,
                capCT$c7,
                capCT$c8,
                capCT$c9)
                
model=crm(marked_dfCT, time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7))
model
model=cjs.hessian(model)
model
```

CJS: AIC: 108.7434, Phi.(Intercept) -0.07818557, p.(Intercept)   -0.99078187


looking at sex and species. phi= apparent survival estimate. p=capture probability (percents) N=population estimate

```{r}
str(marked_dfCT)
marked_dfCT$sex.fac <- as.factor(marked_dfCT$sex)
marked_dfCT$species.fac <- as.factor(marked_dfCT$species)
modCT <- crm(marked_dfCT, 
        model="JS", 
        groups = c("sex.fac"), 
        model.parameters=list(Phi = list(formula = ~sex.fac), 
                              p = list(formula = ~1)))
modCT$results$beta #found online but not sure how to use it
modCT

cjs.hessian(modCT)

str(modCT)
# get values on real scale and not logit or log scales
modCT$results$reals


modCT <- crm(marked_dfCT, 
        model="JS", 
        groups = c("species.fac"), 
        model.parameters=list(Phi = list(formula = ~species.fac), 
                              p = list(formula = ~species.fac)))
modCT$results$beta 
modCT
modCT$results$reals
```

JS sex.fac: AIC  :  212.9627, N  estimate= 37.04946. Species.fac: AIC  :  283.2178, $N estimate= 7.948358e-15



# Run with RMark (need to download Program MARK in order for this to work). This is looking at CT only

```{r}
MarkPath='C:/Program Files (x86)/MARK'
mark(marked_dfCT, model.parameters=list(p=list(formula=~1)),output=FALSE)$results$beta

mark(marked_dfCT, model = "Jolly", model.parameters=list(p=list(formula=~1)),output=FALSE)$results$beta

mark(marked_dfCT, model.parameters=list(Phi=sex.fac),output=FALSE)$results$beta


```

 AICc { Phi(~1)p(~1) } = 108.91003, Phi:(Intercept) -0.078064, p:(Intercept)   -0.990952



CJS and JS could not be run for any other mark/recap site because there were not enough recaptures.



