---
title: "caphistory2"
author: "Bethany Liberto"
date: "June 16, 2019"
output:
  word_document: default
editor_options:
  chunk_output_type: inline
---

Using CJS and JS. JS -> for population estimates. CJS -> for time-dependent survival and recapture rates of a given species 

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = T,
                      #results = "hide") ##This just hides outputs when knitting because it was like 60 pages
options(warn=-1) ##This hides warnings, again, just really for knitting. Can be turned back on with warn=0
```
  
  
```{r}
library(marked)
library(RMark)
library(ggplot2)
```


abundance/effort AIC vvvv

```{r}

data <- read.csv("Dragonfly1718Abundance.csv", stringsAsFactors = F)

### Each possible combination of variables needs to be analyzed.  The lowest score (not counting the full model?) wins. #####
mod1<-glm(data$adjusted.abundance~data$pH)
mod2<-glm(data$adjusted.abundance~data$elevation)
mod3<-glm(data$adjusted.abundance~data$fishpresence)
mod4<-glm(data$adjusted.abundance~data$lenticlotic)  
mod5<-glm(data$adjusted.abundance~data$openwater)
mod6<-glm(data$adjusted.abundance~data$pH*data$elevation)
mod7<-glm(data$adjusted.abundance~data$pH*data$elevation*data$fishpresence)
mod8<-glm(data$adjusted.abundance~data$pH*data$elevation*data$fishpresence*data$lenticlotic)
mod9<-glm(data$adjusted.abundance~data$pH*data$elevation*data$fishpresence*data$lenticlotic*data$openwater) 
mod10<-glm(data$adjusted.abundance~data$elevation*data$fishpresence)
mod11<-glm(data$adjusted.abundance~data$elevation*data$fishpresence*data$lenticlotic)
mod12<-glm(data$adjusted.abundance~data$elevation*data$fishpresence*data$lenticlotic*data$openwater)
mod13<-glm(data$adjusted.abundance~data$fishpresence*data$lenticlotic)
mod14<-glm(data$adjusted.abundance~data$fishpresence*data$openwater) #### Chicken dinner  #####
mod15<-glm(data$adjusted.abundance~data$fishpresence*data$lenticlotic*data$openwater) #### Almost the same ####
mod16<-glm(data$adjusted.abundance~data$lenticlotic*data$openwater)



AIC(mod1, k=2)  ### k = 2 is the "penalty" for adding more factors #####
AIC(mod2, k=2)
AIC(mod3, k=2)
AIC(mod4, k=2) 
AIC(mod5, k=2)
AIC(mod6, k=2)
AIC(mod7, k=2)
AIC(mod8, k=2)
AIC(mod9, k=2) 
AIC(mod10, k=2)
AIC(mod11, k=2)
AIC(mod12, k=2)
AIC(mod13, k=2)
AIC(mod14, k=2) #### Chicken dinner  #####
AIC(mod15, k=2) #### Almost the same, 0.1258 difference ####
AIC(mod16, k=2)


```




richness (just to see) vvvvv

```{r}
### Each possible combination of variables needs to be analyzed.  The lowest score (not counting the full model?) wins. #####
mod1<-glm(data$Richness~data$pH)
mod2<-glm(data$Richness~data$elevation)
mod3<-glm(data$Richness~data$fishpresence)
mod4<-glm(data$Richness~data$lenticlotic)  #### Chicken dinner  #####
mod5<-glm(data$Richness~data$openwater)
mod6<-glm(data$Richness~data$pH*data$elevation)
mod7<-glm(data$Richness~data$pH*data$elevation*data$fishpresence)
mod8<-glm(data$Richness~data$pH*data$elevation*data$fishpresence*data$lenticlotic)
mod9<-glm(data$Richness~data$pH*data$elevation*data$fishpresence*data$lenticlotic*data$openwater)
mod10<-glm(data$Richness~data$elevation*data$fishpresence)
mod11<-glm(data$Richness~data$elevation*data$fishpresence*data$lenticlotic)
mod12<-glm(data$Richness~data$elevation*data$fishpresence*data$lenticlotic*data$openwater)
mod13<-glm(data$Richness~data$fishpresence*data$lenticlotic)
mod14<-glm(data$Richness~data$fishpresence*data$openwater)
mod15<-glm(data$Richness~data$fishpresence*data$lenticlotic*data$openwater)
mod16<-glm(data$Richness~data$lenticlotic*data$openwater)



AIC(mod1, k=2)  ### k = 2 is the "penalty" for adding more factors #####
AIC(mod2, k=2)
AIC(mod3, k=2)
AIC(mod4, k=2) #### Chicken dinner  #####
AIC(mod5, k=2)
AIC(mod6, k=2)
AIC(mod7, k=2)
AIC(mod8, k=2)
AIC(mod9, k=2)
AIC(mod10, k=2)
AIC(mod11, k=2)
AIC(mod12, k=2)
AIC(mod13, k=2)
AIC(mod14, k=2)
AIC(mod15, k=2)
AIC(mod16, k=2)

```



these are all the JS and CJS for each location. Starting with NMBS
```{r}
capNMBS <- read.csv("capturehistoryNMBS.csv", stringsAsFactors = F)
str(capNMBS)
print(capNMBS)
summary(capNMBS)
capNMBS <- capNMBS[which(!is.na(capNMBS$c1)), ]
summary(capNMBS)

library(reshape)
## Format data for MARK

# get data formatted for MARK/RMark
capNMBS$ch <- paste(capNMBS$c1,
                capNMBS$c2,
                capNMBS$c3,
                capNMBS$c4,
                capNMBS$c5,
                capNMBS$c6,
                capNMBS$c7,
                capNMBS$c8,
                capNMBS$c9,
                capNMBS$c10,
                capNMBS$c11,
                capNMBS$c12,
                " ",
                ";",
                sep="")


data=data.frame(capNMBS)
data
transform=melt(capNMBS, id.vars="mark.number")
pivot=cast(transform, mark.number ~ value)
pivot[is.na(pivot)]=0

pivot[,2:ncol(pivot)][pivot[,2:ncol(pivot)] != 0] = 1
lh <- 11;


pivot$eh <- apply(pivot[2:lh],1,paste,collapse="") # concatenates encounter columns into eh
pivot[2:lh] <- NULL # drops individual encounter columns
# create commented tag
pivot$mark.number <- paste("/*", pivot$mark.number, "*/", sep=" ")
# sort by descending encounter histories
pivot <- pivot[order(data$ch,decreasing=TRUE),]
# tack on the frequency for the individual
pivot$end <- "1;";
# write out the input file
write.table(pivot,file="cjs-pivot.inp",sep=" ",quote=F,col.names=F,row.names=F);

pivot

```



```{r}
## Format data for marked package

marked_dfNMBS <- capNMBS

marked_dfNMBS$ch <- paste0(capNMBS$c1,
                capNMBS$c2,
                capNMBS$c3,
                capNMBS$c4,
                capNMBS$c5,
                capNMBS$c6,
                capNMBS$c7,
                capNMBS$c8,
                capNMBS$c9,
                capNMBS$c10,
                capNMBS$c11,
                capNMBS$c12)
                
marked_dfNMBS$sex.fac <- as.factor(marked_dfNMBS$sex)
marked_dfNMBS$species.fac <- as.factor(marked_dfNMBS$species)

model1=crm(marked_dfNMBS,  
          time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
          groups = c("sex.fac"), 
          model.parameters=list(Phi = list(formula = ~sex.fac)),
          hessian=TRUE)

model1
model=cjs.hessian(model1)
model1
model1$results$reals

model2=crm(marked_dfNMBS,  
          time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
          groups = c("sex.fac"), 
          model.parameters=list(p = list(formula = ~sex.fac)),
          hessian=TRUE)

model2
model=cjs.hessian(model2)
model2
model2$results$reals

model22=crm(marked_dfNMBS,  
          time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
          groups = c("sex.fac"), 
          model.parameters=list(Phi = list(formula = ~1), p= list(formula = ~1)),
          hessian=TRUE)

model22
model=cjs.hessian(model22)
model22$results$reals

model3=crm(marked_dfNMBS,  
          time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
          groups = c("species.fac"), 
          model.parameters=list(Phi = list(formula = ~species.fac)),
          hessian=TRUE)

model3
model=cjs.hessian(model3)
model3
model3$results$reals

model4=crm(marked_dfNMBS,  
          time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
          groups = c("species.fac"), 
          model.parameters=list(p = list(formula = ~species.fac)),
          hessian=TRUE)


model4
model=cjs.hessian(model4)
model4
model4$results$reals

model23=crm(marked_dfNMBS,  
          time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
          groups = c("species.fac"), 
          model.parameters=list(p = list(formula = ~1), Phi = list(formula = ~1)),
          hessian=TRUE)


model23
model=cjs.hessian(model23)
model23
model23$results$reals

```

CJS: AIC  :  184.6449, Phi.(Intercept) -0.4564347, p.(Intercept)   -1.5498460

looking at sex and species. phi= apparent survival estimate. p=capture probability (percents) N=population estimate

```{r}
str(marked_dfNMBS)
marked_dfNMBS$sex.fac <- as.factor(marked_dfNMBS$sex)
marked_dfNMBS$species.fac <- as.factor(marked_dfNMBS$species)
#modNMBS <- crm(marked_dfNMBS, 
        #model="JS", 
        #groups = c("sex.fac"), 
        #model.parameters=list(Phi = list(formula = ~sex.fac), 
                              #p = list(formula = ~1)))

modNMBS1 <- crm(marked_dfNMBS, 
        model="JS", 
        groups = c("sex.fac"), 
        time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
        model.parameters=list(N = list(formula = ~sex.fac)),
        hessian=TRUE)

modNMBS1$results$beta #found online but not sure how to use it
modNMBS1

cjs.hessian(modNMBS1)

str(modNMBS1)
# get values on real scale and not logit or log scales
modNMBS1$results$reals


#modNMBS <- crm(marked_dfNMBS, 
       # model="JS", 
       # groups = c("species.fac"), 
      #  model.parameters=list(Phi = list(formula = ~species.fac),    #should have for one or the other, not both
                           #   p = list(formula = ~species.fac)))

modNMBS2 <- crm(marked_dfNMBS, 
                model="JS", 
                groups = c("species.fac"), 
                time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
                model.parameters=list(N = list(formula = ~species.fac)),
                hessian=TRUE)

modNMBS2$results$beta 
modNMBS2
modNMBS2$results$reals
cjs.hessian(modNMBS2)

str(modNMBS2)
# get values on real scale and not logit or log scales
modNMBS2$results$reals

modNMBS3 <- crm(marked_dfNMBS, 
                model="JS", 
                groups = c("species.fac"), 
                time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
                model.parameters=list(N = list(formula = ~species.fac)),
                hessian=TRUE)

modNMBS3$results$beta 
modNMBS3
modNMBS3$results$reals

cjs.hessian(modNMBS3)

str(modNMBS3)
# get values on real scale and not logit or log scales
modNMBS3$results$reals

```

```{r}
#NMBS.table <- collect.models(type = "Jolly")

#This is supposed to get all the models into a nice table so I can compare AIC values but it isn't working (because it only works with RMark)

#####This is my reminder to find a way to make a table of my models AIC values#########
```



# Run with RMark (need to download Program MARK in order for this to work). This is looking at NMBS only

```{r}
#MarkPath='C:/Program Files (x86)/MARK'

#data(marked_dfCNMBS)
#p.sex = list(formula=~marked_dfNMBS$sex.fac)
#Phi.ct= list(formula=~1)
#p.ct = list(formula=~1)
#Phi.sex = list(formula=~marked_dfNMBS$sex.fac)

#mod5=mark(marked_dfNMBS, model = "CJS", model.parameters=list(p=p.ct, Phi=Phi.ct), output=FALSE, time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9))$results$beta

#mod6=mark(marked_dfNMBS, model = "CJS", model.parameters=list(p=p.sex, Phi=Phi.ct), output=FALSE, time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9))$results$beta

#mark(marked_dfNMBS, model.parameters=list(p=list(formula=~species.fac)),output=FALSE, time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9))$results$beta

#mark(marked_dfNMBS, model = "Jolly", model.parameters=list(p=list(formula=~1)),output=FALSE,time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9), options="SIMANNEAL")$results$beta
#This Jolly code doesn't run properly. 

```


####okay now for CT

```{r}
capCT <- read.csv("capturehistoryCT.csv", stringsAsFactors = F)
str(capCT)
print(capCT)
summary(capCT)
capCT <- capCT[which(!is.na(capCT$c1)), ]
summary(capCT)

library(reshape)
## Format data for MARK

# get data formatted for MARK/RMark
capCT$ch <- paste(capCT$c1,
                capCT$c2,
                capCT$c3,
                capCT$c4,
                capCT$c5,
                capCT$c6,
                capCT$c7,
                capCT$c8,
                capCT$c9,
                " ",
                ";",
                sep="")


data=data.frame(capCT)
data
transform=melt(capCT, id.vars="mark.number")
pivot=cast(transform, mark.number ~ value)
pivot[is.na(pivot)]=0

pivot[,2:ncol(pivot)][pivot[,2:ncol(pivot)] != 0] = 1
lh <- 11;


pivot$eh <- apply(pivot[2:lh],1,paste,collapse="") # concatenates encounter columns into eh
pivot[2:lh] <- NULL # drops individual encounter columns
# create commented tag
pivot$mark.number <- paste("/*", pivot$mark.number, "*/", sep=" ")
# sort by descending encounter histories
pivot <- pivot[order(data$ch,decreasing=TRUE),]
# tack on the frequency for the individual
pivot$end <- "1;";
# write out the input file
write.table(pivot,file="cjs-pivot.inp",sep=" ",quote=F,col.names=F,row.names=F);

pivot

```



```{r}
## Format data for marked package

marked_dfCT <- capCT

marked_dfCT$ch <- paste0(capCT$c1,
                capCT$c2,
                capCT$c3,
                capCT$c4,
                capCT$c5,
                capCT$c6,
                capCT$c7,
                capCT$c8,
                capCT$c9)
             
#model=crm(marked_dfCT, time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7))
#model
#model=cjs.hessian(model)
#model

marked_dfCT$sex.fac <- as.factor(marked_dfCT$sex)
marked_dfCT$species.fac <- as.factor(marked_dfCT$species)

model5=crm(marked_dfCT,  
          time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
          groups = c("sex.fac"), 
          model.parameters=list(Phi = list(formula = ~sex.fac)),
          hessian=TRUE)

model5
model=cjs.hessian(model5)
model5
model5$results$reals

model6=crm(marked_dfCT,  
          time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
          groups = c("sex.fac"), 
          model.parameters=list(p = list(formula = ~sex.fac)),
          hessian=TRUE)

model6
model=cjs.hessian(model6)
model6
model6$results$reals

model25=crm(marked_dfCT,  
          time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
          groups = c("sex.fac"), 
          model.parameters=list(p = list(formula = ~1), Phi = list(formula = ~1)),
          hessian=TRUE)

model25
model=cjs.hessian(model25)
model25
model25$results$reals

model7=crm(marked_dfCT,  
          time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
          groups = c("species.fac"), 
          model.parameters=list(Phi = list(formula = ~species.fac)),
          hessian=TRUE)

model7
model=cjs.hessian(model7)
model7
model7$results$reals

model8=crm(marked_dfCT,  
          time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
          groups = c("species.fac"), 
          model.parameters=list(p = list(formula = ~species.fac)),
          hessian=TRUE)


model8
model=cjs.hessian(model8)
model8
model8$results$reals

model26=crm(marked_dfCT,  
          time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
          groups = c("species.fac"), 
          model.parameters=list(p = list(formula = ~1), Phi= list(formula = ~1)),
          hessian=TRUE)


model26
model=cjs.hessian(model26)
model26
model26$results$reals
```



looking at sex and species. phi= apparent survival estimate. p=capture probability (percents) N=population estimate

```{r}
str(marked_dfCT)
marked_dfCT$sex.fac <- as.factor(marked_dfCT$sex)
marked_dfCT$species.fac <- as.factor(marked_dfCT$species)
#modCT <- crm(marked_dfCT, 
        #model="JS", 
        #groups = c("sex.fac"), 
        #model.parameters=list(Phi = list(formula = ~sex.fac), 
                              #p = list(formula = ~1)))

modCT1 <- crm(marked_dfCT, 
        model="JS", 
        groups = c("sex.fac"), 
        time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
        model.parameters=list(N = list(formula = ~sex.fac)),
        hessian=TRUE)

modCT1$results$beta 
modCT1

cjs.hessian(modCT1)

str(modCT1)
# get values on real scale and not logit or log scales
modCT1$results$reals


#modCT <- crm(marked_dfCT, 
        #model="JS", 
        #groups = c("species.fac"), 
        #model.parameters=list(Phi = list(formula = ~species.fac), 
                              #p = list(formula = ~species.fac)))

modCT2 <- crm(marked_dfCT, 
        model="JS", 
        groups = c("species.fac"), 
        time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
        model.parameters=list(N = list(formula = ~species.fac)),
        hessian=TRUE)

modCT2$results$beta 
modCT2
modCT2$results$reals

modCT3 <- crm(marked_dfCT, 
        model="JS", 
        groups = c("species.fac"), 
        time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7),
        model.parameters=list(N = list(formula = ~1)),
        hessian=TRUE)

modCT3$results$beta 
modCT3
modCT3$results$reals
```




# Run with RMark (need to download Program MARK in order for this to work). This is looking at CT only

```{r}
#MarkPath='C:/Program Files (x86)/MARK'

#data(marked_dfCT)
#p.sex = list(formula=~marked_dfCT$sex.fac)
#Phi.ct= list(formula=~1)
#p.ct = list(formula=~1)
#Phi.sex = list(formula=~marked_dfCT$sex.fac)

#mod1=mark(marked_dfCT, model = "CJS", model.parameters=list(p=p.ct, Phi=Phi.ct), time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7), output=FALSE)$results$beta
#mod2=mark(marked_dfCT, model = "CJS", model.parameters=list(p=p.sex, Phi=Phi.ct), time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7), output=FALSE)$results$beta
#mod3=mark(marked_dfCT, model = "CJS", model.parameters=list(p=p.ct, Phi=Phi.sex), time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7), output=FALSE)$results$beta
#mod4=mark(marked_dfCT, model = "CJS", model.parameters=list(p=p.sex, Phi=Phi.sex), time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7), output=FALSE)$results$beta

#mod2=mark(marked_dfCT, model = "Jolly", model.parameters=list(N=list(formula=~1)), time.intervals = c(7, 8, 6, 7, 4, 9, 3, 7), output=FALSE, options="SIMANNEAL")$results$beta

#THIS RUNS REALLY SLOW. I read in the package doc that Jolly often doesn't run at all, which is why there is the options part in this code. Without it, it won't run at all. With it, it runs, but takes forever and doesn't ever fully converge 
#####

#mark(marked_dfCT, model.parameters=list(Phi=sex.fac),output=FALSE)$results$beta

```




```{r}
#??Rmark

#CT.table <- collect.models(type = "CJS")

```




For just crimsons at NMBS
```{r}

capC <- read.csv("capturehistoryCrimsonsNMBS.csv", stringsAsFactors = F)
str(capC)
print(capC)
summary(capC)
capC <- capC[which(!is.na(capC$c1)), ]
summary(capC)

library(reshape)
## Format data for MARK

# get data formatted for MARK/RMark
capC$ch <- paste(capC$c1,
                capC$c2,
                capC$c3,
                capC$c4,
                capC$c5,
                capC$c6,
                capC$c7,
                capC$c8,
                capC$c9,
                " ",
                ";",
                sep="")


data=data.frame(capC)
data
transform=melt(capC, id.vars="mark.number")
pivot=cast(transform, mark.number ~ value)
pivot[is.na(pivot)]=0

pivot[,2:ncol(pivot)][pivot[,2:ncol(pivot)] != 0] = 1
lh <- 11;


pivot$eh <- apply(pivot[2:lh],1,paste,collapse="") # concatenates encounter columns into eh
pivot[2:lh] <- NULL # drops individual encounter columns
# create commented tag
pivot$mark.number <- paste("/*", pivot$mark.number, "*/", sep=" ")
# sort by descending encounter histories
pivot <- pivot[order(data$ch,decreasing=TRUE),]
# tack on the frequency for the individual
pivot$end <- "1;";
# write out the input file
write.table(pivot,file="cjs-pivot.inp",sep=" ",quote=F,col.names=F,row.names=F);

pivot

```



```{r}
## Format data for marked package

marked_dfC <- capC

marked_dfC$ch <- paste0(capC$c1,
                capC$c2,
                capC$c3,
                capC$c4,
                capC$c5,
                capC$c6,
                capC$c7,
                capC$c8,
                capC$c9)
             
marked_dfC$sex.fac <- as.factor(marked_dfC$sex)
marked_dfC$species.fac <- as.factor(marked_dfC$species)

#CJS
modelC1=crm(marked_dfC,  
          time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
          groups = c("sex.fac"), 
          model.parameters=list(Phi = list(formula = ~sex.fac)),
          hessian=TRUE)

modelC1
model=cjs.hessian(modelC1)
modelC1
modelC1$results$reals

#JS

modC2 <- crm(marked_dfC, 
        model="JS", 
        groups = c("sex.fac"), 
        time.intervals = c(23, 4, 9, 2, 4, 15, 9, 1, 2, 7, 9),
        model.parameters=list(N = list(formula = ~sex.fac)),
        hessian=TRUE)

```


CJS and JS could not be run for any other mark/recap site because there were not enough recaptures and I'm not even going to waste my time right now.



